The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-07-13 VegOzone.tex
Addition to abstract - not spell checked.

\begin{abstract}
  The summer 2018 and summer 2019 did not differ significantly in the level of ozone. On the other hand, the temperatures (precipitation and cloud cover/PAR?) differed quite a lot. We also observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019. This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years. Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons. Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.  
\end{abstract}

>>>>

\begin{abstract}
  The past two decades saw a steady reduction in episodic high peak concentrations of ozone over Europe and North America -- a sucessful result of air quality regulations. At the same time, tropospheric background ozone levels kept increasing in the northern hemisphere until recently. An increased occurance of persitant heatwaves, as one possible future scenario, generates favourable conditions for the formation of ambient ozone. Vegetation, highly specialized to sub-artic climate, is vulnerable to rapid environmental changes due to global warming and might become more susceptible to ozone induced damage. The summer of 2018 can, hence, be regarded as a good test case for future scenarios. In both 2018 and 2019, we have monitored ambient ozone concentrations at the Svanhovd research station in Northern Norway. Ambient ozone concentrations did not differ significantly between the two years. On the other hand, temperatures, precipitation, and irradiance differed significantly to some extend. At the same time, we have observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.

  This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years. Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons. Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.  
\end{abstract}
